---
title: "Data Wrangling Lecture Code and Questions"
output: html_document
editor_options: 
  chunk_output_type: console
---

Adapted from UO Psych R Bootcamp: <https://uopsych-r-bootcamp-2020.netlify.app/post/07-dplyr/>

## Instructions

- Open the visual markdown editor.
- For the "Your Turn" questions, you'll need to un-comment the code, and fill in the blanks.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

------------------------------------------------------------------------

**Filter Example 1**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55)
```

**Filter Example 2**

```{r}
penguins %>% 
  filter(species %in% c("Adelie", "Gentoo"),
         island %in% c("Dream", "Torgersen"))
```

------------------------------------------------------------------------

**Arrange Example 1**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(body_mass_g)
```

**Arrange Example 2**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(desc(body_mass_g))
```

------------------------------------------------------------------------

## Your Turn 1 (Completed for you)

1.  Import the file `pragmatic_scales_data.csv` (located in the `data` directory) using the `read_csv` function. Save it to an object named `ps_data`. **Completed for you. Just run as is.**

```{r}
# Q1. 
ps_data <- read_csv("data/pragmatic_scales_data.csv")
```

2.  Filter rows for cases in the "No Label" `condition` and arrange the resulting observations by descending order of `age`.

```{r}
#Q2. 

# ps_data %>% 
#   filter(___) %>% 
#   arrange(___)
```

3.  Select observations from the "No Label" condition for kids 3 years old or younger.

```{r}
# Q3. 

# ps_data %>% 
#   filter(___)
```

4.  Select observations for kids between the ages of 3 and 4. (Extra challenge: look up the documentation for `between()` and use this function.)

```{r}
# Q4. 

# ___ <- ps_data %>% 
#   filter(___)
```

------------------------------------------------------------------------

**Select Example 1**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(body_mass_g) %>% 
  select(species:bill_depth_mm)
```

**Select Example 2**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(body_mass_g) %>% 
  select(species, starts_with("bill_"))
```

**Select Example 3**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(body_mass_g) %>% 
  select(ends_with("_mm"))
```

------------------------------------------------------------------------

**Mutate Example 1**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(body_mass_g) %>% 
  select(starts_with("bill_")) %>% 
  mutate(bill_length_m = bill_length_mm/1000)
```

**Mutate Example 2**

```{r}
penguins %>% 
  filter(species == "Gentoo" & bill_length_mm > 55) %>% 
  arrange(body_mass_g) %>% 
  select(starts_with("bill_")) %>% 
  mutate(bill_length_mm = as.character(bill_length_mm))
```

------------------------------------------------------------------------

## Your Turn 2

1.  In `ps_data`, select only the variables `age` and `condition`.

```{r}
# Q1.

# ps_data %>% 
#   select(___)
```

2.  As we did with indexing in base R, you can use the minus sign (`-`) to "de-select" columns. Keep everything in `ps_data` except `subid` and `condition`.

```{r}
# Q2.

# ps_data %>% 
#   select(___)
```

3.  Select the columns `correct` and `condition` without naming them, using their positions or de-selecting other variables.

```{r}
# Q3. 

# ps_data %>% 
#   select(___)
```

4.  Use `mutate()` to convert `condition` from type "character" to type "factor". Hint: what function could you pipe to directly to check this?

```{r}
#Q4. 

# ps_data %>% 
#   mutate(___) %>% 
#  ___
```

------------------------------------------------------------------------

**Summarize Example 1**

```{r}
penguins %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            max_flipper_length = max(flipper_length_mm, na.rm = TRUE))
```

------------------------------------------------------------------------

**Summarize (pt 2) Example 1**

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))
```

**Summarize (pt 2) Example 2**

```{r}
penguins %>% 
  group_by(species, island) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))
```

------------------------------------------------------------------------

# Your Turn 3

1.  From `ps_data`, get the total number of correct trials for each subject and call this variable `num_correct`. Hint: you can use the summary function `sum()`.

```{r}
# Q1. 

# ps_data %>% 
#   group_by(___) %>% 
#   summarize(___ = sum(___))
```

2.  Now get the total number of correct trials (`num_trials`) for each unique combination of `condition` and `item` and arrange the resulting rows by descending order of `num_correct`. Which combination of condition and item had the most/least correct responses?

```{r}
# Q2. 

# ps_data %>% 
#   group_by(___, ___) %>% 
#   summarize(___ = sum(___)) %>% 
#   arrange(desc(___))
```

3.  Lastly, calculate the proportion of correct responses for each `condition` and call this variable `prop_correct` (hint: because `correct` is coded as 0 and 1, you can do this by taking the mean of `correct`). In the same pipeline, create a bar plot that shows the differences between the mean proportion of correct responses between the two conditions, and color the bars by condition.

```{r}
# Q3. 

# ps_data %>% 
#   group_by(___) %>% 
#   summarize(prop_correct = mean()) %>% 
#   ggplot(___(x = ___, y = ___, fill = ___)) + 
#   ___() +
#   coord_flip() # don't need to change this
```

------------------------------------------------------------------------

## `sample_n` and `sample_frac`

```{r}
penguins %>% sample_n(5, replace = FALSE)
```

```{r}
penguins %>% sample_frac(0.1, replace = FALSE)
```
